Probability of causal contact between interstellar civilizations through MonteCarlo simulations

### 2020-05-07 submit_v03/manuscript.tex
otro parrafo...

In what follows, we describe the experimental setup chosen to estimate
the probabilities of causal contacts and several derived quantities in
terms of three independent parameters, namely, the mean time span
between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$,
their mean lifetime $\left<\tau_s\right>$, and the maximum distance a
signal can be detected by another \ceti{} ($D_{max}$).
%
Intuitively, the shorter the $\tau_a$ parameter and the larger the
$\tau_s$ and $D_{max}$ parameters are, the greater the probability of
the existence of causal contacts between pairs of \cetis{} would be.
%
We also need to propose theoretical distribution functions for both
the distribution of lifespans ($\tau_s$) and the distribution of the
number of \cetis{} per unit time \citep{maccone_evolution_2014,
Sotos_biotechnology_2019}, related to the time span between the
appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$
is shorter, it produces a greater density of \cetis{}).
%
The shapes of these distributions are set to a fixed law, as discussed
in Sec.~\ref{SS_PDF_shape}.

@@

\ffn{1} We illustrate in the Fig.~\ref{F_scheme} the schematic
representation of the region causally connected to a given \ceti{}.
%
To this end, we use space--time diagrams, where time is represented on
the horizontal axis, and space is represented in the vertical axis.
%
For the sake of simplicity, we show in the plot the light travel
distance, i.e., the distance traversed by any signal spreading from
the emitter at light speed (for example, any electromagnetic signal).
%
In this scheme, a light pulse would follow a trajectory represented by
a line at 45 degrees from the axes, given the units of the time and
space axes are years and light years, respectively.
%
Panel (a) of the Figure represents the region $R$ (shaded polygon) in
space and time which is causally connected to the emitter (left
vertex), following an event (hereafter A type event) in which the node
acquires the communication capability, or becomes ''active''.
%
The sphere of first contact (SFC) is a sphere centered on the emitter,
represented in the Figure by the left angle of the region $R$.
%
This sphere grows until its radius reaches the D$_{max}$ distance, in
which it would be no longer detectable due to the decrease in the
energy per unit area, which falls under the assumed detectability
threshold.
%
Similarly, the surface of last contact (SLC) is another sphere that
grows from an event in which the nodes ceases to possess the
communication capability (D type event), and carries the last signal
produced by the emitter.
%
The region which is causally connected to the emitter is then limited
by these two spheres, and has the shape of a sphere (before the D
event) or of a spherical shell (after the D event).
%
The temporal intervals for the communication between two \cetis{} are
represented in the panel (b) of the Figure.
%
In this scheme, the entire lifetime of a \ceti{} can then be
represented by a polygon, limited in the spatial axis by the maximum
distance of the signal $D_{max}$ and in the time axis by the time span
between the starting time (A event, concave vertex) and the ending
time (D event, convex vertex).
%
This representation allows to easily visualize the important events
that result from the existence and communications of two \cetis{}.
%
The points in time where the \cetis{} acquire their communicating
capacity are dubbed ''Awakening'' events ($A_1$ and $A_2$ in the
Figure).
%
Similarly, the points in time where the \cetis{} loss their
communicating capacity are dubbed ''Doomsday'' events ($D_1$ and $D_2$
in the Figure).
%
The points in space--time where the first contact is produced for each
one of the \cetis{}, are defined as ''Contact'' events, shown as $C_1$
and $C_2$ in the Figure.
%
Finally, the two points where the contact is lost for each one the
\cetis{}, are denominated ''Blackout'' events, shown as $B_1$ and
$B_2$ in the Figure.
%
The receiver node E$_1$ can listen signals from the emitter node
E$_2$, from the first contact event (t=C$_{12}$) up to the last
contact event (t=B$_{12}$).
%
Similarly, the receiver node E$_2$ can listen signals from the emitter
node E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
%
The time intervals for the open communication channel are determined
by the ''C'' and ''B'' type events.
%
It is important to notice that there is a time delay between the
contact events of the two \cetis{} involved in this analysis, and also
a time delay between the two blackout events, so that the time range
when a bidirectional contact is possible occurs between the maximum
time of the contact points and the minimum time between the blackout
points.

@@

The temporal structure that emerges naturally from the experimental
setup implies that,
as a fundamental property of the simulations,
a causal connection can be produced without the need of the two
involved locations being concurrent or active at the same time.
%
This property of the system arises as a
consequence of the large spatial and temporal scales, where a message
is transmitted at the (relatively small) light speed.
%
Although a \ceti{} could be active for enough time so as to transmit a
message at large distances, the limited power of the message and the
dilution that runs as the squared distance from the source, imposes a
detectability limit.
%
As a consequence of this limitation and of their finite lifetime,
considered as the period of time between the acquisition and loss of
communicating capacity, each \ceti{} will fill a spherical shell
region of the galaxy, limited by two concentric spherical surfaces.
%
The leading front, or surface of first contact (SFC) grows from the
central \ceti{} until it reaches the maximum distance $D_{max}$.
%
Thus, the volume reached causally by a \ceti{} is initially a growing
sphere.
%
Following the end of the civilization, there is still a region which
is filled with the emitted signals.
%
This approach has been also considered other statistical models
\citep[e.g., ][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
%
The trailing front, or surface of last contact (SLC) also grows from
the central \ceti{}, with a delay with respect to the SFC equivalent
to the lifetime of the node, and produces a spherical shell region.
%
Any other \ceti{} within this region will be aware of the originating
node, even if it has disappeared before the time of contact.
%
This region will grow if the surface of first contact has not yet
reached the maximum distance $D_{max}$, and will shrink otherwise
until the surface of last contact reaches $D_{max}$, producing as a
result the loss of all signal from the central \ceti{}.
%
In our approach, we consider a model galaxy where the width of the
disk is negligible with respect to the radius of the disk.
%
In the 2D simulation only the intersection of the communicating
spherical shells with the plane of the galaxy is relevant, and produce
the corresponding circles or rings for the filled spheres or annular
regions of the spherical shells, respectively.
%
The initially growing communicating sphere, is shown over space--time
diagrams, where time is represented on the horizontal direction, and
space is represented in the vertical direction.
%
In the Fig.~\ref{F_scheme} the two
emitters in panel (b), E$_1$ and E$_2$, reach each other at different times.
%
The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
%
Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
%
The type and length of causal contact in both directions depend on the
distance and time lag between the awakening events, the maximum
distance that a signal can reach and the time period in which each
emitter is active.
%
\ffn{2}
%
In the Fig.~\ref{F_messages} we show a schematic representation of the
possible cases in which a \ceti{} E can be in causal contact with a
\ceti{} R.
%
The duration of the causal contact in one direction depends on several
factors, mainly the lifetimes, the time period between the A events
and the distance between the two nodes involved in the causal
contact.
%
As it can be seen, the direction of the causal contact does not imply
any ordering for the A events of the involved nodes.
%
In the implementation of the DE simulation, the six cases shown in the
Figure determine the succession of events.
%
Each time an event is produced, the list of upcoming events is updated
according to these cases.

>>>>

In what follows, we outline the experimental setup adopted to estimate the probabilities of causal contacts and several derived quantities.
%
This is done in terms of three independent parameters, namely, the mean time span between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$, their mean lifetime $\left<\tau_s\right>$, and the maximum distance a signal can be detected by another \ceti{} ($D_{max}$).
%
Intuitively, the probability of the existence of causal contacts between pairs of \cetis{} would be larger for smaller $\tau_a$ parameters, higher $\tau_s$ or higher $D_{max}$ parameter values.
%
We also propose theoretical distribution functions for both the lifespans ($\tau_s$) and the number of \cetis{} per unit time \citep{maccone_evolution_2014, Sotos_biotechnology_2019}.
%
The later is related to the time span between the appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$ is shorter, it produces a greater density of \cetis{}).
%
The analytical expresiones for these distributions are set to a fixed law, as discussed in Sec.~\ref{SS_PDF_shape}.

@@

\ffn{1} Space—time diagrams, where time and space are represented on
the horizontal and vertical axes, respectively, are suitable tools to
describe the causal connections among different nodes.
%
We illustrate in the Fig.~\ref{F_scheme} the schematic representation
of the region causally connected to a given \ceti{}.
%
For the sake of simplicity, we show in the plot the light travel
distance, i.e., the distance traversed by any signal spreading from
the emitter at light speed (for example, any electromagnetic signal).
%
In this scheme, a light pulse would follow a trajectory represented by
a line at 45 degrees from the axes, given the units of the time and
space axes are years and light years, respectively.
%
Panel (a) of the Figure represents the region $R$ (shaded polygon)
which is causally connected to the emitter (left vertex).
%
This region develops after an event (hereafter dubbed A-type event) in
which the node acquires the communication capability, or becomes
''active''.
%
The sphere of first contact (SFC) is centred on the emitter,
represented in the Figure by the left angle of the region $R$.
%
This sphere grows until its radius reaches the D$_{max}$ distance, in
which it would be no longer detectable due to the decrease in the
energy per unit area, which falls under the assumed detectability
threshold.
%
Similarly, the surface of last contact (SLC) is another sphere that
grows from an event in which the nodes ceases to possess the
communication capability (D type event), and carries the last signal
produced by the emitter.
%
The region which is causally connected to the emitter is then limited
by these two spheres and therefore has the shape of a filled sphere
(before the D event) or of a spherical shell (after the D event).
%
The temporal intervals for the communication between two \cetis{} are
represented in the panel (b) of the Figure.
%
In this scheme, the entire lifetime of a \ceti{} can be represented by
a polygon, as shown in the Figure.
%
The polygon is limited in the spatial axis by the maximum distance of
the signal $D_{max}$ and in the time axis by the time span between the
A-type event (concave vertex) and the D-type event (convex vertex).
%
This representation allows to visualize the important events that
result from the presence and communications of two \cetis{}.
%
The points in time where the \cetis{} acquire their communicating
capacity are dubbed ‘’Awakening’’ events ($A_1$ and $A_2$ in the
Figure).
%
Similarly, the instants in time where the \cetis{} lose their
communicating capacity are dubbed ‘’Doomsday’’ events ($D_1$ and $D_2$
in the Figure).
%
The points in space--time where the first contact is produced for each
one of the \cetis{}, are defined as ''Contact'' events, shown as $C_1$
and $C_2$.
%
Finally, the two points where the contact is lost for each one the
\cetis{}, are denominated ''Blackout'' events, presented as $B_1$ and
$B_2$ in the Figure.
%
The receiver node E$_1$ can listen signals from the emitter node
E$_2$, from the first contact event (t=C$_{12}$) until the last
one (t=B$_{12}$).
%
Similarly, the receiver node E$_2$ can listen signals from the emitter
node E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
%
The time intervals for the open communication channel are determined
by the ''C'' and ''B'' type events.
%
It is noteworthy the fact that there is a time delay between the
contact events of the two \cetis{} involved in this analysis, and also
a time delay between the two blackout events.
%
Therefore, the time range when a bidirectional contact is possible
occurs between the maximum time of the contact points and the minimum
time between the blackout points.

@@

The temporal structure that emerges from the experimental setup implies that, as a fundamental property of the simulations, a causal connection can be produced without 
requiring that the two nodes are active at the same time.
%
This property of the system arises as a consequence of the large spatial and temporal scales, where a message is transmitted at the (relatively small) light speed.
%
Although a \ceti{} could be active for a large enough period to transmit a message at large distances, the limited power of the message and the dilution that depends on the squared distance from the source imposes a detectability limit.
%
As a consequence of this limitation and of their finite lifetime, considered as the period between the acquisition and loss of communicating capacity, each \ceti{} will fill a spherical shell region of the galaxy, limited by two concentric spherical surfaces.
%
The leading front, or surface of first contact (SFC) grows from the central \ceti{} until it reaches the maximum distance $D_{max}$.
%
Following the end of the civilization, there is still a region which is filled with the emitted signals.
%
This approach has been also considered in other statistical models \citep[e.g., ][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
%
The trailing front, or surface of last contact (SLC) also grows from the central \ceti{}, with a delay with respect to the SFC equivalent to the lifetime of the node, and produces a spherical shell region.
%
Any other \ceti{} within this region will be aware of the originating node, even if it has disappeared before the time of contact.
%
This region will grow if the surface of first contact has not yet reached the maximum distance $D_{max}$, and will shrink after a D-type event until the surface of last contact reaches $D_{max}$, producing as a result the loss of all signal from the central \ceti{}.
%
In our approach, we consider a model galaxy where the width of the disk is negligible with respect to the radius of the disk.
%
In the 2D simulation only the intersection of the communicating spherical shells with the plane of the galaxy is relevant, and produce the corresponding circles or rings for the filled spheres or annular regions of the spherical shells, respectively.
%
The initially growing communicating sphere is shown over space--time diagrams, where time is represented on the horizontal direction, and space is represented in the vertical direction.
%
In the Fig.~\ref{F_scheme} the two emitters in panel (b), E$_1$ and E$_2$, reach each other at different times.
%
The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
%
Emitter $i$ can listen to emiter $j$ between C$_{ij}$ and B$_{ij}$.
%
The type and length of causal contact in both directions depend on the distance and time lag between the awakening events, the maximum distance that a signal can reach and the time period in which each emitter is active.

\ffn{2}
%
In the Fig.~\ref{F_messages} we show a schematic representation of the possible cases in which a \ceti{} E can be in causal contact with a \ceti{} R.
%
The duration of the causal contact in one direction depends on several factors, mainly the lifetimes, the time period between the A events and the distance between the two nodes involved in the causal contact.
%
As it can be seen, the direction of the causal contact does not imply any ordering for the A events of the involved nodes.
%
In the implementation of the DE simulation, the six cases shown in the Figure determine the succession of events.
%
Each time an event is produced, the list of upcoming events is updated according to these cases.
